---
title: 'Wk 9 : NY Times Movie Reviews API'
author: "Aaron Grzasko"
date: "October 30, 2016"
output: 
    html_document:
        theme: simplex
        highlight: haddock
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, warning=FALSE, message=FALSE)
```

## Assignment Overview  
  
In this assignment, I use R to interface with the NY Times movie reviews API.  
  
The main goal is to transform the JSON output from the NY Times into an R, data frame format.  
  
## Resources  
I used the following resources to complete this assignment:  
  
* [NY Times Movie Reviews documentation](http://developer.nytimes.com/movie_reviews_v2.json): This page provides documentation on the basic mechanics of the movie reviews API.  

* [NY Times Github](https://github.com/NYTimes/public_api_specs/blob/master/movie_reviews/movie_reviews_v2.md)  This document contains supplemental information for using the NY Times movie reviews API.  

* [Vignette on fetching JSON data in R](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-apis.html) The example code from this page relies exclusively on the `fromJSON()` function in the jsonlite` package to read in the JSON format data.  
  

## Load Libaries  
```{r}
if (!require(jsonlite)) install.packages('jsonlite')
if (!require(knitr)) install.packages('knitr')
if (!require(DT)) install.packages('DT')
```
  
## API Key and URL for NY Times API 
  
While access to the  NY Times API is free, one must first first register [here](http://developer.nytimes.com/signup) to retrieve an API key. This key is required for accessing the data.    
  
Refer to the scripts below.  The `api_val` object stores the API key.  
  
The `base_url` object stores the URL for the NY Times Movie reviews API.    
  
```{r}
# api key 
api_val <- "9963c662bb2d444cb55470a17bceacd9"

# url of nyt movie reviews api
base_url <- "http://api.nytimes.com/svc/movies/v2/reviews/search.json?"
```


## Manipulating the API URL String  
  
The NY Times API documentation provides helpful tips on how to query the underlying data.  
  
Specifically, the documentation outlines a series of parameters that can be added to the API's base URL for refining search criteria.  
  
These parameters include:  
  
* *query*: used to match keywords.  When querying, users enter "+" to separate words instead of spaces.  
* *critics-pick*: "Y" or "N".  Used to distinguish movies that have been "highlighted" by NYT.  
* *reviewer*: used to limit results to a specific movie critic
* *publication-date*: this is the date the movie review was originally published in NYT.  Format = "YYYY-MM-DD".  Users can specify both start and end dates by separating the two date values with a semicolon.  
* *opening-date*:  the date the movie first opened in NYC.  Format = "YYYY-MM-DD".  Users can specify both start and end dates by separating the two date values with a semicolon.  
* *offset*:  used for extracting multiple pages of records.  The value specified must be whole number multiple of 20, as the API will extract 20 records at a time.  To specify records 21-40, for instance, the user would enter `offset = 20`.  
* *order*:  this parameter is used to sort the displayed records.  
  

## Function for Generating Modified URL Strings 
  
I created a custom function, `string_append()`, that produces the appropriate URL, given a user's input for various parameter values.  My intent with this function was to partially automate the process of querying the API.  

```{r}
# url of ny_times


string_append <- function(base_url, api_val, query_val,critic_val, reviewer_val, pub_val, opendt_val,offset_val, order_val ){
    
    api_string <- base_url 
    query_cd <- 'query='
    critic_cd <- "critics-pick=" 
    pub_cd <- "publication-date="
    opendt_cd <- "opening-date="
    offset_cd <- "offset="
    reviewer_cd <- "reviewer="
    order_cd <- "order="
    api_cd <- "&api-key="
    
    count_parem <- 0
    
    if (!missing(query_val)){
        count_parem <- count_parem + 1
        api_string <- paste0(api_string, query_cd,query_val)
    }
    if(!missing(critic_val)){
        ifelse(count_parem == 0, api_string <- paste0(api_string, critic_cd, critic_val), 
            api_string <- paste0(api_string,"&",critic_cd,critic_val))
        count_parem <- count_parem + 1
    }
    if(!missing(reviewer_val)){
        ifelse(count_parem == 0, api_string <- paste0(api_string, reviewer_cd, reviewer_val), 
            api_string <- paste0(api_string,"&",reviewer_cd, reviewer_val))
        count_parem <- count_parem + 1
    }
    
    if(!missing(pub_val)){
        ifelse(count_parem == 0,  api_string <- paste0(api_string, pub_cd, pub_val),
            api_string <- paste0(api_string,"&",pub_cd, pub_val))
        count_parem <- count_parem + 1
    } 
    if(!missing(opendt_val)){
        ifelse(count_parem == 0,  api_string <- paste0(api_string, opendt_cd, opendt_val), 
            api_string <- paste0(api_string,"&",opendt_cd, opendt_val))
        count_parem <- count_parem + 1
    } 
    if(!missing(offset_val)){
        ifelse(count_parem == 0,  api_string <- paste0(api_string, offset_cd, offset_val),
            api_string <- paste0(api_string,"&",offset_cd, offset_val))
        count_parem <- count_parem + 1
    } 
    if(!missing(order_val)){
        ifelse(count_parem == 0,  api_string <- paste0(api_string, order_cd, order_val),
            api_string <- paste0(api_string,"&",order_cd, order_val))
        count_parem <- count_parem + 1
    }

    api_string <- paste0(api_string,api_cd, api_val)
    
}

 
```

## Example 1:  Ghostbusters Movies  
  
Let's query the data, using specified parameter values in the `string_append()` function. 

I use the the `fromJSON()` function in the jsonlite package to store the initial data to a data frame object.  
  
In this example, I am querying any reviews with the keyword "ghostbuster".  
  

```{r}
# search criteria:  keyword = "ghostbuster"

url_string <- string_append(base_url = base_url, api_val = api_val, query_val = "ghostbuster")

# store data as data frame
raw <- fromJSON(url_string)

# unmodified results
results <- raw$results

# check names of columns 
colnames(results)


# data scrubbing, and subsetting fields 

results_redux <- subset(results, select = c("display_title","mpaa_rating","critics_pick","publication_date","opening_date","date_updated","link"))


results_redux$link_url <- results_redux$link[[2]]
results_redux <- subset(results_redux, select = c("display_title","mpaa_rating","critics_pick","publication_date","opening_date","date_updated","link_url"))

# print results
kable(results_redux)

```
  
Let's confirm that our data is stored as a data frame:  
  
```{r}
class(results_redux)
```

## Example 2: 2016 Critics' Picks  
  
In this second example, I perform a slightly more complicated search, using the following criteria:  
  

* opendt_val = "2016-01-01":  movies released 1/1/2016 or later.  
* critic_val = "Y": movies distinguished as "critics' picks".  
* offset_val:  using values of 0, 20, and 40 in a for loop to pull 3 pages (or 50 entries) of data.  
* order_val:  ordering by movie release date  
  
  
```{r}

# initialize empty data frame
df <- data.frame(display_title = character(),mpaa_rating = character(), critics_pick = integer(),publication_date=character(),opening_date = character(),date_updated = character(),link_url = character())

# loop through first 3 pages of results
for (i in 0:2){
    
    url_string <- string_append(base_url = base_url, api_val = api_val, opendt_val = "2016-01-01", critic_val = "Y", offset_val = i*20,  
        order_val="by-opening-date")
    raw <- fromJSON(url_string)
    result <- raw$result
    result_redux <- subset(result, 
        select = c("display_title","mpaa_rating","critics_pick","publication_date","opening_date","date_updated","link"))

    # light data scrubbing
    result_redux$link_url <- result_redux$link[[2]]
    result_redux <- subset(result_redux, 
        select = c("display_title","mpaa_rating","critics_pick","publication_date","opening_date","date_updated","link_url"))    

    df <- rbind(result_redux, df)
}

# print results

datatable(df)
```
  
  
Once again, we'll confirm that our data is stored in a data frame:  
```{r}
class(df)
```
